---
bibliography: references.bib
---
## Appendix Graphics

*Linear Regression Model*
```{r}
library(tidyverse); theme_set(theme_bw())
reg_data_long <- read_csv(".\\Data\\reg_data_long.csv")
# head(reg_data_long)
reg_data_long$female_prop <- reg_data_long$FemalePop/reg_data_long$TotalPop

set.seed(123)

training_pct <- .8
Z <- sample(nrow(reg_data_long), floor(training_pct*nrow(reg_data_long)))

train <- reg_data_long[Z,]
# dim(train)
test <- reg_data_long[-Z,]
# dim(test)

test$County <- factor(test$County, levels = levels(train$County))

full_model <- lm(DeathsperThou ~ TotalPop + MalePop + FemalePop + MedianAge + MaleMedianAge + FemaleMedianAge + Under5yoPop + Under18yoPop + Pop21andOlder + Pop55andOlder + Pop60andOlder + Pop65andOlder + num_desig + amer_indian + black + hispanic + asian + multi_race + pac_island + white + poverty + labor_force + unemployed + median_income + TotalDeaths + TotalCases + prop_hpsa + nonwhite_prop + Rating + Deaths + unempl_prop + poverty_prop + prop55older + Cases + HospNumber + CasesperThou + female_prop , data = train)
```

NA Counts for Regression Data
```{r}
na_counts <- colSums(is.na(reg_data_long))
na_counts

```

Linear Fit of Variables
```{r}

library(car)
avPlots(full_model)

```
Here, we can see that some variables provide a better linear fit than others. For example, variables such as `MaleMedianAge` and `FemaleMedianAge` both are more or less scattered along the regression line. However, other variables such as `MalePop` and `FemalePop` do not fit the line well. Instead, the data is gathered in one spot and has almost no linear shape. These graphs give some reason for concern that the basic full model will not be the strongest. Not all of the variables are linear so a reduced model might be a better fit. 


## Group Member Accomplishments

*Roland Abi*


*Elise Buellesbach*

Found and proposed the original CA data set (`California Hospital Inpatient Mortality Rates and Quality Ratings`). Completed the linear regression models (both full and reduced). Wrote the executive summary, summary/conclusion, ethical analysis, and references. 

*Amanda Concepcion*


*Spencer Grewe*

Acquired CA demographic data (`data4`). Cleaned datasets 1, 3, and 4. Contributed toward tidying and joining datasets 1, 3, and 4 for the regression tasks. Carried out 627 tasks involving model selection for regression. 
