---
bibliography: references.bib
---
## Appendix Graphics

*Linear Regression Model*
```{r}
library(tidyverse); theme_set(theme_bw())
reg_data_long <- read_csv(".\\Data\\reg_data_long.csv")
# head(reg_data_long)
reg_data_long$female_prop <- reg_data_long$FemalePop/reg_data_long$TotalPop

set.seed(123)

training_pct <- .8
Z <- sample(nrow(reg_data_long), floor(training_pct*nrow(reg_data_long)))

train <- reg_data_long[Z,]
# dim(train)
test <- reg_data_long[-Z,]
# dim(test)

test$County <- factor(test$County, levels = levels(train$County))

full_model <- lm(DeathsperThou ~ TotalPop + MalePop + FemalePop + MedianAge + MaleMedianAge + FemaleMedianAge + Under5yoPop + Under18yoPop + Pop21andOlder + Pop55andOlder + Pop60andOlder + Pop65andOlder + num_desig + amer_indian + black + hispanic + asian + multi_race + pac_island + white + poverty + labor_force + unemployed + median_income + TotalDeaths + TotalCases + prop_hpsa + nonwhite_prop + Rating + Deaths + unempl_prop + poverty_prop + prop55older + Cases + HospNumber + CasesperThou + female_prop , data = train)
```

NA Counts for Regression Data
```{r}
na_counts <- colSums(is.na(reg_data_long))
na_counts

```
Plotting Full Model
```{r}
plot(full_model)
```
These plots give some insight to the shape of the model. For example, the plotting on the QQ-plot does not demonstrate a strong linear relationship. The edges of the plot deviate from the line suggesting that the model is not perfectly linear. Further, the scattered dots on the Leverage plot show several points on the line of Cook's Distance signifying that they have overstated influence on the model. With these concerns in mind, I want to investigate the linearity of the model further. Does a linear model make sense? To get a better understanding of the data, I created some basic plots to demonstrate the shape of the data and get an idea of if a linear model will make sense here ([source on code](https://www.statology.org/plot-multiple-linear-regression-in-r/))

Linear Fit of Variables
```{r}

library(car)
avPlots(full_model)

```
Here, we can see that some variables provide a better linear fit than others. For example, variables such as `MaleMedianAge` and `FemaleMedianAge` both are more or less scattered along the regression line. However, other variables such as `MalePop` and `FemalePop` do not fit the line well. Instead, the data is gathered in one spot and has almost no linear shape. These graphs give some reason for concern that the basic full model will not be the strongest. Not all of the variables are linear so a reduced model might be a better fit. 


## Group Member Accomplishments

*Roland Abi*
Developed the research question for the classification, and found the data set for `california health professional shortage area` and cleaned and transformed the data. Completed logistic regression model (full and reduced), as well as cross validation approaches contributed to completing decision tree and boosting models, carried out 627 tasks for model selection and two wrote the classification section of the summary as well as future direction.   

*Elise Buellesbach*

Found and proposed the original CA data set (`California Hospital Inpatient Mortality Rates and Quality Ratings`). Completed the linear regression models (both full and reduced). Wrote the executive summary, summary/conclusion, ethical analysis, and references. 

*Amanda Concepcion*


*Spencer Grewe*

Acquired CA demographic data (`data4`). Contributed toward cleaning, tidying and joining data sets 1, 3, and 4 for the regression tasks. Carried out 627 tasks involving model selection for regression. 


## Statement of Data Availability

The source code, data, and full supporting information are available at the [shared repository](https://github.com/sbgrewe/STAT627-FinalProj), which will be made public on the date of the submission of this report. We declare that all authors equally own the contents of the repository. 
