
```{r}
library(randomForest)
library(tidyverse); theme_set(theme_bw())
reg_data_long <- read_csv(".\\..\\..\\Data\\reg_data_long.csv")
# head(reg_data_long)
reg_data_long$Rating <- as.factor(reg_data_long$Rating)
reg_data_long$female_prop <- reg_data_long$FemalePop/reg_data_long$TotalPop

set.seed(123)
Z <- sample(1:nrow(reg_data_long), 0.8*nrow(reg_data_long), replace = FALSE)
train <- as.data.frame(reg_data_long[Z,])
test <- as.data.frame(reg_data_long[-Z,])
```

There are nine unique predictors in the data.

```{r}
p <- 9 
RF <- OptimalTrees <- Yhat <- RMSEP <- vector(mode = "double", length = p)

set.seed(123)
for (k in 1:p) {
  RF <- randomForest(DeathsperThou ~ Rating*CasesperThou + prop_hpsa + nonwhite_prop + median_income + unempl_prop 
								 + poverty_prop + Rating*prop55older + Rating*HospNumber,
								 data = train, mtry = k)
  OptimalTrees[k] <- which.min(RF$mse)
  RF <- randomForest(DeathsperThou ~ Rating*CasesperThou + prop_hpsa + nonwhite_prop + median_income + unempl_prop 
								 + poverty_prop + Rating*prop55older + Rating*HospNumber,
								 data = train, mtry = k, ntree = OptimalTrees[k])
  Yhat <- predict(RF, newdata = test)
  RMSEP[k] <- sqrt(mean((Yhat - test$DeathsperThou)^2) )
}

which.min(RMSEP)
OptimalTrees[7]
```

Cross validation produces the optimal number of predictors and number of trees for the training data: 7 predictors and 398 trees respectively. 

```{r}
set.seed(123)
model_RF <- randomForest(DeathsperThou ~ Rating*CasesperThou + prop_hpsa + nonwhite_prop + median_income + unempl_prop + poverty_prop + Rating*prop55older + Rating*HospNumber, data = train, mtry = 7, ntree = 398)
model_RF
plot(model_RF)

Y_hat <- predict(model_RF, newdata = test)
mean((test$DeathsperThou-Y_hat)^2)
```

The random forest model has a prediction MSE of 0.02992, which corresponds to a 37.64% decrease in prediction MSE with respect to LASSO regression. We find that

```{r}
varImpPlot(model_RF)
```

We find that the dominant predictor, as expected, was the number of heart attack cases per thousand, though now the rating of the hospital again appears tp 

